Multi-level modelling with Openflexo/FML - A contribution to the \mpc

### 2021-05-25 3-Introduction.tex

In order to tackle this problem, the \emph{multi-level} modeling paradigm have been introduced. Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as a flexibilization of the relations between levels.

This approach is gaining interest lately within the modeling community as shown by the MULTI level-modeling workshop and the daghtul seminar~\citep{almeida2018multi} and number of tools. Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created. This paper is a response to the latest multi-level modeling challenge which consist of modeling processes. We do so by using model federation and the OpenFlexo framework.

Model federation is a modeling approach in which loosely couple reified models. Links among models can be within a same level of abstraction or across levels of abstractions which gives, as our answer to the challenge shows, a great flexibility.

Our solution, which is fully implemented and executable, meets all the requirements.

The rest of the paper is organized as follows.

>>>>

In order to tackle this problem, the \emph{multi-level} modeling paradigm have been introduced. Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as a flexibilization of the relations between them. This paradigm is gaining traction lately within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools and the celebration of topic specific workshops\footnote{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/WebHome} and seminars~\citep{almeida2018multi}. Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created. This paper is a response to the latest multi-level modeling challenge
which consists in providing a solution to the problem of specifying and enacting processes. Solutions to the \emph{Multi-level Process Challenge} must fulfill a number of requirements for a process representation defined at an abstract process-definition level and at various more concrete domain-specific levels, resulting in a multi-level hierarchy of related models. We provide a solution based on model federation~\citep{Golra2016-federation}.

Model federation is a multi model management approach based on the use of virtual models and loosely coupled links. The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models and links serve as control components used to present different views to the users and maintain synchronization. Our solution is based on this model federation infrastructure, namely, the use of virtual models and links. Concretely, we use OpenFlexo and its internal Flexo Modeling Language (FML), a language to create, link and manage virtual models, in order to achieve our goal. Our solution, which is fully implemented and executable, meets all the requirements.

%Links among models can be within a same level of abstraction or across levels of abstractions which gives,

%as our answer to the challenge shows, a great flexibility.

The rest of the paper is organized as follows. Section~\ref{sec:technology} presents our approach and the technology it relies on. Section~\ref{sec:analysis} is the analysis of the problem. We detail our model in section~\ref{sec:model} and show it satisfies the challenge requirements in section~\ref{sec:requirements}. We assess the modeling solution in section~\ref{sec:discussion}. Section~\ref{sec:relatedwork} presents the related work before the conclusion~\ref{sec:conclusion}.

### 2021-05-25 5-Analysis.tex
Updates from Overleaf

Like any modeling approaches, these different steps could also be achieved in any order. But our goal during the challenge requirement analysis remains to lead to define a FML Virtual Models architecture as a federation of virtualmodels.

Also as our understanding, the use case analysis leads to identify two main modeling axis (as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}):
\begin{itemize}
    \item the horizontal axis is characterized as the an ontological instantiation axis, in the sense that the domain type definition (ProcessType as example) is referenced by an instance definition (Process, a.e). In our base  VirtualModel, each domain type definition is referenced by its instance definition, as developed in the section \ref{sec:ProcessEnactmen}

@@

any disambiguations of the  case  description and assumptions made, any potentially added requirements

quelles hypothèses supplémentaires ?

+ architecture de la modélisation (ce qui se trouve dans le ppt), le détail se retrouvera dans la section suivante

>>>>

Like any modeling approaches, these different steps could also be achieved in any order. But our goal during the challenge requirement analysis remains to lead to define a FML Virtual Models architecture as a federation of virtualmodels. As shown in figure~\ref{fig:MultilevelArchitecture}, we developped two metamodels (Base and ACME processes abstractions) instantiatied by two models (XSure and ACME processes). This follows the way the challenge is presented and this organisation allows for the possibility of easy extensions.

Also as our understanding, the use case analysis leads to identify two main modeling axis (as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}):
\begin{itemize}
    \item the horizontal axis is characterized as the ontological instantiation axis, in the sense that the domain type definition (ProcessType as example) is referenced by an instance definition (Process, a.e). In our base  VirtualModel, each domain type definition is referenced by its instance definition, as developed in the section \ref{sec:ProcessEnactment}

@@

%\emph{any disambiguations of the  case  description and assumptions made, any potentially added requirements}

\todo[inline]{quelles hypothèses supplémentaires ?}
\todo[inline]{+ architecture de la modélisation (ce qui se trouve dans le ppt), le détail se retrouvera dans la section suivante}
